Stochastic Optimal Control of Single Neuron Spike Trains\\ --\\ Reply to Referees at the JNE

### 2014-02-18 floats.tex
Adding ML Note

\documentclass[12pt]{iopart}
%\documentclass{article}

\usepackage{amsmath}
\usepackage{amscd}
\usepackage{amssymb} 
\usepackage{amsfonts}     
\usepackage{amsthm}
\usepackage{amsfonts} 
\usepackage{amsthm}

\usepackage{tikz}
\usetikzlibrary{arrows,snakes,backgrounds}
% \usepackage{circuitikz}
% \usepackage{pgf}
% \usetikz
\usepackage{subfig}

\usepackage{algpseudocode}
\usepackage{algorithm}

\usepackage[super]{nth}
\usepackage{appendix}
% \usepackage{listings}
\usepackage{color}

\usepackage{hyperref}
%\usepackage{url}

\usepackage{cleveref}
\usepackage{cancel}
\usepackage{slashbox}

\usepackage{aviolov_style}
\usepackage{local_style}

\usepackage{iopams}

% \newtheorem{thm}{Theorem}[section]
% \newtheorem{lemma}{Theorem}[thm]
% \theoremstyle{definition}
% \newtheorem{ex}{Example}[thm]
% \newtheorem{defn}{Definition}[thm]

% \def\includegraphic{}
% \def\includegraphics{}

\begin{document}

\title{Stochastic Optimal Control of Single Neuron Spike Trains}
\author{Alexandre Iolov$^{1,2}$, Susanne Ditlevsen$^2$, Andr\'e Longtin$^{1,3}$ 
		}	
\address{%
     (1) Department of Mathematics and Statistics, University of Ottawa, Ottawa,
    Canada
\\
    (2) Department of Mathematical Sciences, University of Copenhagen,
Copenhagen, Denmark
\\
    (3) Department of Physics, University of Ottawa,
    Ottawa, Canada
}
\ead{aiolo040@uottawa.ca}			
% \address{Department of Mathematics and Statistics, University of Ottawa, Ottawa, Canada 
% \\ Department of Mathematical Sciences, University of Copenhagen, Copenhagen,
% Denmark }
% \ead{$<$\href{mailto:aiolo040@uottawa.ca}
% 		{aiolo040 at uottawa dot ca}$>$}

% \date{\today}

\begin{abstract}

{\em Objective.} External control of exact spike times in single neurons is
important for basic understanding of neuronal activity, and has potential in
brain-machine interfaces, for improving neural prostheses, and in medical
applications related to neuronal disorders, such as Parkinson's disease or
epilepsy. The goal of this paper is the design of optimal controlled electrical
stimulation patterns to a cell to achieve a target spike train under
physiological constraints not to damage tissue. {\em Approach.} We pose a
stochastic optimal control problem to precisely specify the spike times in a
noisy leaky integrate-and-fire model of a neuron. In particular, we allow for
the noise to be of arbitrary intensity. The optimal control problem is solved
using Dynamic Programming if the controller has access to the voltage
(closed-loop control) and using a Maximum Principle for the transition density
if the controller only has access to the spike times (open-loop control). {\em
Main results.} We have developed a stochastic optimal control scheme to obtain
precise spike times, which is not limited by the intensity of the noise in the
system and is applicable in both the supra-threshold and sub-threshold regimes.
Given this scope, our simulations show that our algorithms produce the desired results. They further reveal that the numerical demands are non-trivial,
such that the algorithms may require efficiency refinements to achieve real-time
control, in particular the open-loop context is more numerically demanding than
the closed-loop one. Furthermore, we show how wrong the control strategy can
perform if we assume that the noise is negligible. {\em Significance.} More
precise targeting and control of neural activity to achieve optimal performance
in a noisy environment, in which most neurons are embedded, have great potential
in health care. The main contribution is the online feedback control of a noisy
system through modulation of the input, taking into account physiological
constraints on the control, as well as incurred costs to diminish adverse
effects.

\end{abstract}
% \tableofcontents MSC KEYS: 49N90 90C39 93E20 93C20 93C95

\maketitle

% \tableofcontents
% TODO: Remove ToCs (for internal consumption only)

\section{Introduction}
Manipulation of individual neurons through external electrical
stimulation provides a means to
controlling the spiking activity. Applications may arise, for
example, in brain-machine interfaces or in artificial prosthetics.
A natural goal is to make a cell produce a pre-specified spike train,
under certain constraints on the modulation of the input.
We consider this problem in the framework of Stochastic Optimal Control and
give both a feedback solution (closed-loop), when the cell voltage is
explicitly observable, as well as an open-loop solution when only the
occurrence of spikes is observable. Importantly, we allow for an
arbitrary noise intensity.

Both theoretically and in practice related problems have been
addressed in the literature. One objective has been to
obtain either minimum or maximum interspike intervals lengths,
when the input is constrained to be between some prespecified upper
and lower bounds, see e.g. \cite{Lee1994,Lefebvre1987} for a
mathematical treatment, or \cite{Danzl2009,Nabi2012,Wu2009} in a neuronal
context. Another objective is to break a pathological synchronous
firing pattern in clusters of neurons, highly relevant for neurological
disorders such as epilepsy and Parkinson's disease,
\cite{Nabi2013a,Nabi2011}, see also \cite{Feng2007b}.

Our objective, namely targeting exact spike times in single neurons, has been
considered mainly in the open-loop context, and in either absence of or for
small noise. In \cite{Ahmadian2011} they use the Spike Response Model,
\cite{Gerstner2002}, to control output target spike trains and implement their
scheme on pyramidal cells in mouse cortical slices. Their method is numerically
efficient and allows for the simultaneous control of many neurons. However the
efficiency of the numerics strongly relies on the assumption that the diffusion
noise, the value of $\b$ in \cref{eq:X_evolution_uo} below, is small enough, see
the discussion above and below equation 12 in \cite{Ahmadian2011}. They only
work with open-loop control. The difference between the objective in
\cite{Ahmadian2011} and what we consider below is that they maximize the
probability of spiking at some given time $t^*$, whereas we minimize the mean squared
difference between the realized spike time, $T_{sp}$ and the desired, $t^*$.
Moehlis et al. \cite{Moehlis2006} work with a Phase Response Model to obtain
spikes at exact times, while keeping the root mean square of the input to a
minimum. A similar approach is taken in \cite{Dasanayake2011}. They do not
consider noise, though, and only work with open-loop control. In
\cite{Nabi2013}, their methods are implemented on brain slices of pyramidal
neurons of rat hippocampus. While the
Phase Response Curve Model is a parsimonious and effective way to describe a
neuron's response to a stimulus, it is only valid in the supra-threshold regime,
where the unstimulated neuron is periodically spiking. Reference
\cite{Feng2003} investigates the control of the firing times of a
leaky integrate-and-fire neuron by varying the intensity of a noise process that
drives the voltage (there is no other intrinsic noise in the model). 
To obtain a given spike time, they choose the objective of minimizing the
variance of the membrane potential at the desired spike time, while forcing the
mean of the membrane potential at this time point to equal the threshold. This
provides exact solutions since it does not involve first-passage times, but has
the drawback that there is a non-negligible probability that the obtained spike
time will be far from the desired spike time.

Our objective of imposing a certain timing sequence for the spike train using an
externally applied control is obtained in both the closed- and
open-loop setting, and we specifically include the noise in the
calculations of the controls. We restrict the controlled input to stay
within pre-specified bounds, and also include a cost function to
minimize intervention.
Our main contributions are that we specifically allow for a non-negligible
noise component in the neural activity when calculating the control, and
consider both open- and closed-loop controls. In particular we do not
restrict our attention to the autonomously-spiking, supra-threshold regime.

The paper is structured as follows: First we describe the neuron model and
formalize the control objective. Then we describe a feedback-based solution,
which assumes that the controller has detailed access to the voltage trajectory.
Then we relax the observation assumption so that the controller only has access
to the spike times. Finally, we compare the two methods through
simulations against a simple-minded control technique which ignores
the stochastic input to the neuron.

\section{Problem Formulation}
A basic but useful model for the neural membrane potential evolution is the
noisy leaky-integrate-and-fire model:
\begin{equation}
\begin{gathered}
dX(t) = \left(\Iext(t) - \frac{X(t)}{\tc} \right) \intd{t} + \b \intd{W},
\\
X(0) = 0,
\\
X(\ts) = \xth \implies
\begin{cases}
X(\ts^+) = 0. & 
\end{cases}
\end{gathered}
\label{eq:X_evolution_uo}
\end{equation}
Here $X(t)$ represents the membrane electric potential at time $t$, which in
absence of input decays to $0$ with a time constant of $\tc$, $dW$ is a Brownian
motion increment scaled by $\b$ and $\Iext(t)$ is the deterministic external input to the cell. Having last
spiked at time $0$, the potential hits $\xth$ at some random time $\ts$, the
potential resets to $0$ and the process starts all over again. 
We write $\ts^+$ for the limit from the right at $\ts$.
Throughout the paper the threshold is set to one, $\xth = 1$.

Suppose that we have some control over the external current such that it can be
decomposed as
\begin{equation}
\Iext(t) = \m + \a(t),
\label{eq:current_mu_alpha}
\end{equation}
where $\m$ is an uncontrollable, but constant part of the external current and
$\a(t)$ is controllable, i.e.\ it can be chosen to achieve some goal.
A natural goal is to attempt to control the spike time, $\ts$.
That is, how do we choose $\a(t)$ such that $\ts \approx \T$, where $\T$ is the
desired spike-time. A natural optimal control objective to achieve this is the
least squares solution
\begin{equation}
\a(\cdot) = \argmin_{\a(\cdot)} \{\, \Exp[(\ts- \T)^2] \,\},
\label{eq:OC_LS_variance}
\end{equation}
where expectation is taken with respect to the distribution of the trajectories
of $X$.

Often the control has certain constraints. The most common are simple box
constraints:
\begin{equation}
\a(t) \in [\amin, \amax] \quad \forall t.
\label{eq:bound_constraints_alpha}
\end{equation}

% \subsection{Optimal Control Formulation}
In addition to \cref{eq:OC_LS_variance}, we will also add to our objective a
running energy cost based on the control. This regularizes the problem
eliminating the subtleties of singular-control situations and it serves to avoid excessive
control as well as to avoid excessive charge building up on the cell, see
\cite{Ahmadian2011}.

So we seek an optimal control, $\a^*$, that solves
\begin{align}
J[\a(\cdot)] =&
\Exp\left[
\e \int_0^\ts  \a^2(s) \intd{s}
+
(\ts - \T \big)^2 \right]
\label{eq:OC_LS_variance_energy}
\\ \notag
\a^*(\cdot) =& \argmin_{\a(\cdot)} J[\a(\cdot)],
\end{align}
where $\e$ measures how much weight we put on minimizing the energy cost.
If $\e = 0$, then we do not care at all about the expended energy cost.

It will often be the case that $\a$ means either a function or a
value of that function at a particular time. This function could be
random, i.e.\ a stochastic process. Since $\a$ is used to control the behaviour
of $X$, it will be at least measurable with respect to the filtration of $X$,
see Ch.\ 11 in \cite{Oksendal2007} for the details. We will try to make clear
below when we are refering to $\a$ as a stochastic process and when we are merely refering to its particular
value at some time. For example, in \cref{eq:OC_LS_variance_energy}, $\a(\cdot)$
refers to the function, while $\a(s)$ refers to that function's value, possibly
random, at time $s$.

We will consider two control contexts -- {\sl closed-loop} and  {\sl open-loop}
control. In closed-loop control, the value of $X$ at time $t$ is observable
and can be used in determining the control. In open-loop control only the spike
times are observable. In the closed-loop context, we write the control as $$\a =
\a(x,t)$$ to indicate its dependence on $X(t)$ and to express that it will be
updated based on the time-course of $X$. In the open-loop context, we write $$\a
= \a(t)$$ to indicate that $\a(\cdot)$ is decided for all times at time 0. The
techniques used to obtain the optimal controls in the two scenarios will be
different. For the closed-loop scenario we use Dynamic Programing,
see \cite{Fleming1975}, while for the open-loop scenario we use a form of the
Maximum Principle applied to the transition density of the controlled process,
see \cite{Borzi2012}.

Crucially, we assume that the model parameters, $\m,\tc,\b$ in
\cref{eq:X_evolution_uo,eq:current_mu_alpha} are known. This is a strong assumption and will be
discussed later.

\subsection{Parameter Regimes}
Different  parameter regimes can be envisioned given \cref{eq:X_evolution_uo},
depending on whether the noise intensity, $\b$, is
relatively high or low, and whether the external, uncontrollable bias current,
$\m$, induces spikes in the absence of noise or not. 
Spikes will occur in the absence of noise, if and only
if $ \m > 1 / \tc,$ which is called the supra-threshold regime. When $\m
\leq 1/\tc$, the regime is called the sub-threshold regime.
In addition we will investigate two values of $\b$, which we
will call high-noise and low-noise, respectively.
Example values for each parameter regime are given in \cref{tab:regimes}, and we
visualize a single path for each regime in \cref{fig:regime_path_examples}.
\begin{table}
\begin{tabular}{|l||{c}|{c}|}
\hline
\backslashbox{$\m$}{$\b$}
& $1.5$ & $0.3$ \\
\hline
$1.5 / \tc $ &Supra-threshold-High-noise & Supra-threshold-Low-noise \\
\hline
$0.1 / \tc$   &Sub-threshold-High-noise & Sub-threshold-Low-noise \\
\hline
\end{tabular}
\caption{Regime labels and example values. Note that for the numerical
experiments below, we use $\tc = 0.5$}
\label{tab:regimes}
\end{table}
%\usepackage{graphics} is needed for \includegraphics
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=0.99\textwidth]{Figs/PathSimulator/path_T=14_combined.pdf}
  \caption[labelInTOC]{Example Trajectories from \cref{eq:X_evolution_uo}
  using the parameter values from \cref{tab:regimes}. 
  A) Supra-Threshold-Low-Noise 
  B) Supra-Threshold-High-Noise 
  C) Sub-Threshold-Low-Noise
  D) Sub-Threshold-High-Noise. 
  Note the multiple crossings very close together in the high-noise regimes
  in B and D. }
  \label{fig:regime_path_examples}
\end{center}
\end{figure}

\section{Closed-Loop Solution - Dynamic Programing}
We now detail the Dynamic Programing approach to obtaining an optimal feedback
control. In closed-loop, the controller can be continuously updated depending on
the realization of the stochastic process, $X$.

Given a time $t$ and a realized voltage value $X(t) = x$, let $(\ts-t)$ be the
unknown and random remaining time to spike, note that $\ts$ is a random
variable. Given arbitrary $t, x$, our remaining-cost objective, $J[\a(\cdot);
x,t]$, will be
\begin{equation}
J[\a(\cdot); x,t]  =
\Exp\left[
\big((\ts-t) - (\T-t) \big)^2  
+
\e \int_t^\ts  \a^2(s) \intd{s} 
\,\Big|\, X(t) = x
\right],
\label{eq:OC_LS_cost_to_go}
\end{equation}
i.e.\ if time $t$ has elapsed without a spike, we now want to minimize
the difference between $(\ts - t)$ and $(\T-t)$, given the current state $x$.
The mean in \cref{eq:OC_LS_cost_to_go} is taken over the distribution of
hitting times, $\ts$, conditional on $X(t)$ or equivalently over the
distribution of forward trajectories of $X$ starting at $x$ and ending at the
threshold. 
Recall that in the closed-loop scenario, we assume that
the value of $X(t)$ is known to the controller. A similar problem is discussed
analytically at length in the book on optimal control by Peter Whittle,
\cite{Whittle1996}, although there is no discussion there of the numerics
required to solve it.

\subsection{Hamilton-Jacobi-Bellman equation}
The Hamilton-Jacobi-Bellman (HJB) equation, see \cite{Fleming1975,Whittle1996}
or the articles \cite{Danzl2009,Nabi2013a} in a neuroscience context, associated
with the optimal control for \cref{eq:OC_LS_cost_to_go} is obtained as follows.
We introduce the value function, $\v(x,t)$, as the minimum of the remaining-cost objective, i.e.\ of
the cost function between the current time $t$ and the desired spike time
$\T$:
\begin{align}
\v(x,t) =&
 \min_{\a(\cdot)_{s \geq t}}
 \{J[\a(\cdot); x,t] \}
\notag
\\
=&
\min_{\a(\cdot)_{s \geq t}}
\Exp\left[
\big((\ts-t) - (\T-t) \big)^2
+
\e \int_t^\ts  \a^2(s) \intd{s}
\,\Big|\,X(t) = x
\right].
\label{eq:OC_LS_mean_value}
\end{align}
Then $\v$ satisfies the following HJB partial differential equation (PDE):
\begin{equation}
\begin{gathered}
\di_t \v(x,t) + \tfrac{\b^2}{2} \di_x^2 \v(x,t) + \\
\min_{\a(x,t) \in [\amin,
\amax]}\Big\{ \e \a^2 (x,t)+ \left(\m + \a(x,t)-\tfrac{x}{\tc}\right) \di_x
\v(x,t) \Big\} = 0.
\label{eq:OC_LS_mean_HJB}
\end{gathered}
\end{equation}
The special feature of \cref{eq:OC_LS_mean_HJB} in contrast to a generic
parabolic PDE is that it contains an embedded optimisation that depends on the
solution, $\v$. For each $x,t$ in the computational domain,
$\a$ is chosen such as to minimize $\{\e \a^2 + \left(\m + \a -
\tfrac{x}{\tc}\right) \di_x \v \}$. Here we can solve for the optimal control, $\a^*(x,t)$ analytically as:
\begin{align}
\a^*(x,t)  =& \argmin_{\a \in [\amin, \amax]}\big\{\e \a^2 + (\m + \a-\tfrac{x}{\tc}) \di_x\v\big\}
\notag
\\
=&
\min \left(\amax, \max\left(\amin, -\frac{\di_x\v(x,t)}{2\e}\right)\right).
\label{eq:OC_LS_variance_energy_quadratic_control}
\end{align}
With this the HJB PDE becomes
\begin{equation}
\di_t \v(x,t) + \tfrac{\b^2}{2} \di_x^2\v+
\e (\astar(x,t))^2 + (\m + \astar(x,t)-\tfrac{x}{\tc}) \di_x\v
= 0.
\label{eq:OC_LS_mean_HJB_bounded_control}
\end{equation}

We need to consider boundary conditions (BCs) for $\v$. If $X(t) = \xth$ then we
have a spike now and $\ts = t$. Thus $$ \v(\xth,t) = (t-\T)^2.$$ At the
threshold, the value function equals the squared difference between the desired
spike time and the realized one. For $x \ll 0$, we assume that $\v$ is not
significantly affected by the change in $x$, i.e.\ that $$ \di_x \v(\xmin, t) =
0 $$ for some lower boundary $\xmin$. Such a boundary condition will be
justified if we impose artificial reflecting boundaries for the diffusion at
$\xmin$. So we need to determine for which $\xmin$ is a reflecting boundary
consistent with the dynamics. For example, we can take $\xmin$ to be two
standard deviations below the mean of the stationary distribution of the
maximally inhibited process, i.e.\ setting $\a=\amin$ in
\cref{eq:current_mu_alpha}. That is we set $\xmin = \tc(\mu + \amin) - 2 \b / \sqrt{\tc/2}$. We further enforce
that $\xmin \leq -0.5$.
% Ooops, we need $\di_x\v\neq 0$ and here we are setting $\di_x\v= 0$ in one of
% the BCs\ldots oh well.

Note that Dynamic Programing and the HJB equation work backwards. Thus the
evolution of the value function proceeds from the future to the past and we need
some Terminal Condition (TCs) at some point in the future, possibly infinity,
from which to start incrementing $\v$ using the dynamics and the BCs. To
determine TCs for $\v$, our idea is simple: if we reach $\T$ without having
spiked we apply maximum control in the positive direction, i.e.\ 
$$t>\T \implies \a(t) = \amax.$$ Thus: \begin{equation}\v(x,\T) = \Exp
\Big[\trem^2 \,\Big|\, X(\T) = x, \a(t) = \amax \Big].
\label{eq:valuefun_TCs}
\end{equation}
Note that we are making an approximation here
- we are ignoring the energy term, $\eps u^2$, in the objective for
$t>\T$. Naturally, this approximation is ever more accurate for $\eps \ll 1$.
We discuss in more detail the validity of the approximation in
\cref{sec:effect_of_eps}.
Alternatively, we could impose this approximate terminal condition at some $t^+
> \T$.

We will see the quantity on the right hand side of \cref{eq:valuefun_TCs}
repeatedly so we will give it a special name.
\begin{equation}
\Ttwo(x) := \Exp \Big[\trem^2 \,\Big|\, X(\T) = x, \a(t) = \amax \Big].
\label{eq:Trem_squared}
\end{equation}
$\Ttwo$ is the second moment of the remaining time to reach the
threshold starting at $X(\T) = x$ and applying $\amax$ throughout. This quantity
can be found easily for all $x$ in the domain by solving a stationary backward Kolmogorov
equation. This is an extension to the calculation of the first moment of an
exit-time as seen in textbooks such as \cite{Jacobs} and we give the
details in \ref{sec:valuefun_TCs}.

Thus, we restate the HJB equation in its fully specified form:
\begin{equation}
\begin{gathered}
\di_t \v(x,t) + \tfrac{\b^2}{2} \di_x^2\v(x,t)+
\e \a^2(x,t) + \big(\m + \a(x,t) -\tfrac{x}{\tc}\big)\cdot \di_x\v(x,t)
= 0,
\\
\a (x,t) = \min \left(\amax, \max\left(\amin, -\frac{\di_x \v(x,t)}{2\e}\right)
\right).
\\
\begin{cases}
\v(\xth,t) = (t-\T)^2  \quad & \textrm{upper BC}
\\
\di_x \v(\xmin, t)  = 0  \quad &\textrm{lower BC}
\\
\v(x,\T)  =
% \Exp \Big[\trem^2 \,\Big|\, X_\T = x, \a(t) = \amax \Big]
% \Exp[\t^2|x, \amax]  \quad
\Ttwo(x)
\,& \textrm{TC}
\end{cases}
\end{gathered}
\label{eq:OC_LS_HJB_full}
\end{equation}
We are solving $\v(x,t)$ over the domain $[\xmin, \xth] \times [0,\T]$.

\subsection{The numerical method for the HJB equation}
\label{sec:hjb_numerix}
We now have a PDE for $\v$ and an algorithm for computing all the BCs and TCs of
this PDE. It is time to discuss the numerical method for solving
\cref{eq:OC_LS_HJB_full}. Since it is one-dimensional in space, it is straight
forward to apply the standard centred finite difference using the
Crank-Nicholson scheme to step in time, see ch.\ 19.2 in \cite{Press1992}. To
resolve the non-linearity, $(\di_x\v)^2$, in the PDE, we treat it as a mixed
implicit-explicit term $$(\di_x \v(x,t_k))^2 = \di_x \v(x,t_k) \cdot \di_x
\v(x,t_k) \approx \underbrace{\di_x \v(x,t_k)}_{\textrm{implicit}} \cdot
\underbrace{\di_x\v(x,t_{k+1})}_{\textrm{explicit}}.$$ Note that the implicit
term is in the previous time $t_k$ instead of, as is conventional, the next
$t_{k+1}$, because we are solving for $\v$ {\sl backwards} in time from
$t_{k+1}$ to $t_k$ down to  $t_0 = 0$.

The numerical scheme is implemented in Python, using the Scipy/Numpy library,
\cite{Scipy}. For the discretization in time and space, we choose $\Delta x$ and
$\Delta t$ in relation to the parameter regimes. In particular, we take $\Delta
x$ to be less than $\b$ divided by the largest possible absolute value of $(\m +
\a - x/\tc)$ for $\a \in [\amin, \amax], x \in [\xmin, \xth]$; in our examples
this always be equal to $\mu + \amax - \xmin/\tc$. This is an attempt to ensure
that numerically we are in the diffusion-dominated regime. We set $\Delta t$ to
be twice the value of $\Delta x$ divided by the largest possible absolute
value of $(\m + \a - x/\tc)$. For example in the Supra-Threshold Low-Noise regime, we
have $\Delta x = 0.0043$ and $\Delta t = 0.0013$. In all regimes, our
discretization results in point-wise convergence of up to at least three
significant digits for the value function. We consider this to be a sufficiently
fine discretization.

We next demonstrate solutions for the value function and the associated
control law for a parameter set from each regime in \cref{tab:regimes}.

\subsection{Solutions of the HJB equation} 
We set $\T=1.5$ and $\eps = 0.001$, so that the primary focus is the accurate
spiking and energy minimization considerations are secondary.

>>>>

\begin{figure}[htp] 
\begin{center} 
  \includegraphics[width=0.99\textwidth]{Figs/PathSimulator/path_T=14_combined.pdf} 
  \caption[labelInTOC]{Example Trajectories from \cref{eq:X_evolution_uo} 
  using the parameter values from \cref{tab:regimes}.  
  A) Supra-Threshold-Low-Noise  
  B) Supra-Threshold-High-Noise  
  C) Sub-Threshold-Low-Noise 
  D) Sub-Threshold-High-Noise.  
  Note the multiple crossings very close together in the high-noise regimes 
  in B and D. } 
  \label{fig:regime_path_examples} 
\end{center} 
\end{figure} 
\begin{figure}[htp] 
\begin{center} 
  \includegraphics[width=\textwidth]{Figs/HJB/Regimes_vc_cuts.pdf} 
  \caption[labelInTOC]{Snapshots of the value function, $\v$, on the left and on 
  the optimal control, $\a$, on the right for $t$ fixed, at the start (blue), 
  mid-point (green) and end-point (red), for each of the four regimes. 
  The desired spike time is set to $\T=1.5$ assuming the previous spike was at 
  $t=0$, the energy penalty, $\eps = 0.001$. The control bounds are $\a \in 
  [-2,2]$.  
  Note that all red value function curves go to zero at the threshold to satisfy 
  the upper BC in \cref{eq:OC_LS_HJB_full}. 
  Note that the green and blue 
  curves in F) are lying on top of each other for all $x$. 
  \\ 
  A,B) Supra-Threshold-Low-Noise 
  C,D) Supra-Threshold-High-Noise 
  E,F) Sub-Threshold-Low-Noise 
  G,H) Sub-Threshold-High-Noise 
  } 
\label{fig:HJB_4regimes_value_control_cuts} 
\end{center} 
\end{figure} 
\begin{figure}[h!] 
\begin{center} 
\includegraphics[width=0.99\textwidth]{Figs/HJB/Regimes_valuesurf.pdf} 
\caption{The numerical solution for the value function $\v(x,t)$ HJB PDE, 
\cref{eq:OC_LS_HJB_full}, for the four different parameter regimes. 
The desired spike time is set to $\T=1.5$. 
The control bounds are $\a \in [-2,2]$.  
The coloured contours for fixed $t$ plotted in thick correspond to the 
curves plotted in the left panels of \cref{fig:HJB_4regimes_value_control_cuts} 
(blue: $t=0$, green: $t=t*/2$, red: $t=t*$). 
\\ 
A) Supra-Threshold-Low-Noise 
B) Supra-Threshold-High-Noise 
C) Sub-Threshold-Low-Noise  
D) Sub-Threshold-High-Noise  } 
\label{fig:HJB_4regimes_value_surf} 
\end{center} 
\end{figure} 
\begin{figure}[htp] 
\begin{center} 
  \includegraphics[width=\textwidth]{Figs/FP_Adjoint/Regimes_cs_singleplot.pdf} 
  \caption[labelInTOC]{The deterministic optimal controls for each parameter 
  regime as functions of time, $t \in [0, \T]$. 
  The desired spike time is set to $\T=1.5$, the energy penalty, $\eps 
  = 0.001$ and the bounds are $\a \in [-2,2]$. 
  From left-to-right, the control curves correspond to the 
  Sub-Threshold-Low-Noise, Sub-Threshold-High-Noise, Super-Threshold-High-Noise, 
  Sub-Threshold-Low-Noise regimes 
  } 
  \label{fig:FBK_Regimes_cs} 
\end{center}   
\end{figure}    
\begin{figure}[h]   
\begin{center} 
\subfloat[Supra-Threshold Low-Noise]{ 
  \includegraphics[width=\textwidth] 
  {Figs/ControlSimulator/Regimes_SUPT_ln_errors_hist.pdf} 
} 
\\ 
\subfloat[Supra-Threshold High-Noise]{ 
  \includegraphics[width=\textwidth] 
  {Figs/ControlSimulator/Regimes_SUPT_HN_errors_hist.pdf} 
  } 
  \\ 
\subfloat[Sub-Threshold Low-Noise]{ 
  \includegraphics[width=\textwidth] 
  {Figs/ControlSimulator/Regimes_subt_ln_errors_hist.pdf} 
} 
\\ 
\subfloat[Sub-Threshold High-Noise]{ 
  \includegraphics[width=\textwidth] 
  {Figs/ControlSimulator/Regimes_subt_HN_errors_hist.pdf} 
}\\ 
  \caption[labelInTOC]{Histogram of the spike timing error for the 
  deterministic (left) vs. open-loop stochastic (centre) vs. closed-loop 
  stochastic (right) control laws. 
  Recall that $\ts$ is the random realized spike time and $\T$ is the target 
  spike time.   
  This is for the same problem as in 
  \cref{fig:control_trajectories_examples}. 
  We have used $N=10000$ sample paths 
  to form the statistics.} 
  \label{fig:error_histograms_det_vs_openloop_vs_stoch} 
\end{center} 
\end{figure} 
\begin{figure}[h] 
\begin{center} 
\includegraphics[width=0.99\textwidth] 
{Figs/ControlSimulator/Composite_Traj.pdf}  
\caption[]{Three examples for the controlled trajectories using the 
deterministic, open-loop stochastic and closed-loop stochastic control approaches. 
The plots on the left, A,C,E, are obtained using a value of $\eps=0.001$, 
while the plots on the right, B,D,F, use $\eps=0.1$ (for a discussion on the 
effect of $\eps$ see \cref{sec:effect_of_eps}). The black vertical line in the plots indicates the 
desired spike-time, $\T$. The parameter values are $\m, \tc, \b = [0.2, 0.5,  1.5]$ (Sub-Threshold 
High-Noise regime), with $\T = 1.5$.  
The upper plots show the voltage evolution, $X(t)$, while the lower 
plots show the applied control, $\a(t)$. Note that the optimal control, $\a(t)$, obtained from 
the deterministic and from the open-loop controls are the same across all three 
samples.} 
\label{fig:control_trajectories_examples}    
\end{center}   
\end{figure} 
\begin{figure}[htp] 
\begin{center} 
  \includegraphics[width=\textwidth]{Figs/TrainController/SUPT_ln_cl_trains_sim_50.pdf} 
  \caption[ ]{How good the {\sl closed-loop} controller is at hitting a target 
  spike train. In the panel above, the blue curve is the empirical firing 
  rate of $M$ controlled trains.   strategy, the red curve is a 
  smoothed-version of the target train. (See text for details of how they are calculated). In the panel below, the 
  dashed, red lines indicate the target times, $\t_n$, while the blue dots are 
  the realizations of the controlled trains.   
  The parameters from the Super-Threshold, Low-Noise 
  regime are used. The target spike train was generated using the model itself, 
  with parameters from the Super-Threshold, High-Noise regime, without an applied 
  control, $\a=0$. 
  There are $N=16$ target spikes and $M=50$ realizations of the control strategy}  
  \label{fig:targettrain_cl_lownoise} 
\end{center} 
\end{figure}   
\begin{figure}[htp]  
\begin{center} 
  \includegraphics[width=\textwidth]{Figs/TrainController/SUPT_ln_ol_trains_sim_50.pdf} 
  \caption[ ]{How good the {\sl open-loop} controller is at hitting a target 
  spike train. 
  The drawn curves have  the same meaning as in 
  \cref{fig:targettrain_cl_lownoise}. 
  The parameters from the Super-Threshold, Low-Noise regime are used.  
  The target spike train was generated using the model itself, with parameters 
  from the Super-Threshold, High-Noise regime, without an applied control, $\a=0$. 
  There are $N=16$ target spikes and $M=50$ realizations of the control strategy} 
  \label{fig:targettrain_ol_lownoise} 
\end{center}  
\end{figure} 
\begin{figure}[htp] 
\begin{center} 
  \includegraphics[width=\textwidth]{Figs/TrainController/SUPT_HN_cl_trains_sim_50.pdf} 
  \caption[ ]{How good the {\sl closed-loop} controller is at hitting a target 
  spike train.  
  The drawn curves have  the same meaning as in 
  \cref{fig:targettrain_cl_lownoise}. 
  The parameters from the Super-Threshold, High-Noise regime 
  are used. The target spike train was generated using the model itself, with 
  parameters from the Super-Threshold, High-Noise regime, without an applied 
  control, $\a=0$. 
  There are $N=16$ target spikes and $M=50$ realizations of the control strategy} 
  \label{fig:targettrain_cl_highnoise} 
\end{center}  
\end{figure} 
\begin{figure}[htp] 
\begin{center} 
  \includegraphics[width=\textwidth]{Figs/TrainController/SUPT_HN_ol_trains_sim_50.pdf} 
  \caption[]{How good the {\sl open-loop} controller is at hitting a target 
  spike train.  
  The drawn curves have  the same meaning as in 
  \cref{fig:targettrain_cl_lownoise}. 
  The parameters from the Super-Threshold, High-Noise regime 
  are used. The target spike train was generated using the model itself, with 
  parameters from the Super-Threshold, High-Noise regime, without an applied 
  control, $\a=0$. 
  There are $N=16$ target spikes and $M=50$ realizations of the control strategy} 
  \label{fig:targettrain_ol_highnoise}     
\end{center} 
\end{figure} 
\begin{figure}[htp] 
\begin{center} 
  \includegraphics[width=\textwidth]{Figs/TrainController/SUPT_HN_cl_aplus_trains_sim_50.pdf} 
  \caption[ ]{Same as \cref{fig:targettrain_cl_highnoise}, but with the control 
  constraints, $[\amin, \amax]$ increased to $[-4,4]$ instead of $[-2,2]$.} 
  \label{fig:targettrain_cl_highnoise_aplus}     
\end{center} 
\end{figure}  
\begin{figure}[htp] 
\begin{center} 
  \includegraphics[width=\textwidth]{Figs/HJB/Regimes_eps_comparison.pdf} 
  \caption[labelInTOC]{Effect of $\e$ on the closed-loop control. Snapshots of 
  $\a(x,t)$ for $t$ fixed, at the start (blue), mid-point (green) and end-point 
  (red), for each of the four regimes and different values of $\eps$. 
  On the left, we use $\eps = 0.001$ and on the right $\eps = 0.1$. 
  The desired spike time is set to 
  $\T=1.5$ and the bounds are $\a \in [-2,2]$.  
  A,B) 
   Supra-Threshold-Low-Noise  
  C,D)  
   Supra-Threshold-High-Noise  
  E,F) 
   Sub-Threshold-Low-Noise  
  G,H) 
   Sub-Threshold-High-Noise} 
\label{fig:HJB_4regimes_control_different_eps}  
\end{center} 
\end{figure} 
\begin{figure}[htp] 
\begin{center} 
  \includegraphics[width=\textwidth]{Figs/FP_Adjoint/Regimes_eps_comparison.pdf} 
  \caption[labelInTOC]{Effect of $\e$ on the open-loop control. 
  The deterministic time-trajectory of $\a(t)$ for each 
  parameter regime as function of time, $t \in [0, \T]$. 
  On the left, we use $\eps = 0.001$ and on the right $\eps =0.1$. 
  The desired spike time is set to $\T=1.5$ and the bounds are $\a \in [-2,2]$. 
    A,B) 
   Supra-Threshold-Low-Noise 
  C,D)  
   Supra-Threshold-High-Noise 
  E,F) 
   Sub-Threshold-Low-Noise 
  G,H) 
   Sub-Threshold-High-Noise} 
  \label{fig:FBK_Regimes_cs_different_es}  
\end{center} 
\end{figure} 
